### 2014-06-16 kypchak-morphfest-2014/poster.tex
some more Kyrgyz numbers and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54480 72bbbca6-d526-0410-a7d9-f06f51895060

		\htwo{Framework: HFST}
		\begin{itemize}
			\item Reimplements Xerox FST formalisms ({\tt lexc} \& {\tt twol})
			\item Also provides a wrapper around popular free/open-source FST toolkits: SFST, OpenFST, and Foma
		\end{itemize}
	}
	\headerbox{Development effort}{name=devel,below=morphtrans}{
		\begin{itemize}
			\item Kumyk transducer based on Kazakh, Tatar transducers
			\item \tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
		\end{itemize}
		\vspace*{-2.5em}
		\hspace*{-2em}\includegraphics[width=0.9\textwidth,angle=270]{../kipchak-lrec-2013/graphs/kumyk-yoldash-cov}
		\vspace*{-2em}
	}

@@

		\headerbox{Categorisation}{name=lexc,below=exout,column=1}{
		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
			\begin{itemize}
				\item Example:\ morphologically distinct adjective classes (not documented elsewhere)
				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
				\item Our approach:\ categorisation (generates/analyses only correct forms)
			\end{itemize}

@@

			\htwo{Number of stems}
			%\vspace{-1.5em}
			\vspace{-0.5em}
			\begin{center}
			\begin{tabular}{lrrr}
				\toprule
				\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{c}{\textbf{Number of stems}} \\ \cmidrule{2-4}
				& \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
				\midrule
				Noun & 2640 & 2795 & 2568 \\
				Verb & 1470 & 1143 & 386 \\
				Adjective & 754 & 816 & 219 \\
				Proper noun & 5701 & 5361 & 1443 \\
				Adverb & 171 & 177 & 63 \\
				Numeral & 63 & 63 & 44 \\
				Conjunction & 46 & 45 & 13 \\
				Postposition & 50 & 43 & 12 \\
				Pronoun & 32 & 28 & 17 \\
				Determiner & 39 & 34 & 9 \\
				\midrule
				Total: & \hiliteone{11224} & \hiliteone{10737} & \hiliteone{4845} \\
				\bottomrule
			\end{tabular}
			\end{center}
			%\vspace{2em}
			\vspace{-0.5em}

@@

%		\scalebox{0.8}{
%			\begin{tabular}{lll}
%				\toprule
%				\textbf{type} & \textbf{lang} & \textbf{contents} \\ % & \textbf{origin} \\
%				\midrule
%				Encyclopædic & kaz & Wikipedia \\ % wpdump & 20131006 \\
%					& tat & Wikipedia \\ % wpdump & 20130225 \\
%					& kum & — \\ % — & — \\
%				\midrule
%				News & kaz & RFE/RL (azattyq.org) \\ % RFE/RL & azattyq.org 2010 \\
%					& tat & tat.tatar-inform.ru \\ % Татар-информ & tat.tatar-inform.ru 2005-2011 \\
%					& kum & Ёлдаш (yoldash.etnosmi.ru) \\ % Ёлдаш & yoldash.etnosmi.ru \\
%				\midrule
%				Religion & kaz & Quran + Bible \\ % quran + bible & kkitap.net, kuran.kz \\
%					& tat & Quran + New Testament \\ % quran + nt & ibt.org.ru, tanzil.net \\
%					& kum & Genesis + New Testament \\ % genesis + nt & ibt.org.ru \\
%				\bottomrule
%			\end{tabular}
%		}
		\scalebox{0.77}{
			\begin{tabular}{llll}
				\toprule
					& \textbf{Wikipedia} & \textbf{News} & \textbf{Religion} \\
				\midrule
					\texttt{\textbf{kaz}} & Wikipedia & azattyk.org & Quran + Bible \\
					\texttt{\textbf{tat}} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
					\texttt{\textbf{kum}} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
				\bottomrule
			\end{tabular}

@@

		\scalebox{0.89}{
			\begin{tabular}{llrrr}
				\toprule
				 & \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) & \textbf{Amb.} \\
				\midrule
				\multirow{4}{*}{\textbf{Kazakh}} & Wikipedia & 25.6M & 85.61 $\pm$ 1.37 & 2.43 \\
					& News & 3.8M & 92.12 $\pm$ 2.72  & 2.88 \\
					& Religion & 851K & 92.49 $\pm$ 1.66  & 2.63 \\\cmidrule{2-5}
				 (r50547) & Average &  & \hiliteone{90.07} $\pm$ 1.91  & 2.64\\
				\midrule
				\multirow{4}{*}{\textbf{Tatar}} & Wikipedia & 159K & 86.35 $\pm$ 2.17  & 2.24\\
					& News & 5.2M & 89.75 $\pm$ 0.07  & 2.30 \\
					& Religion & 382K & 91.25 $\pm$ 2.55 & 2.24 \\\cmidrule{2-5}
				 (r50260) & Average &  & \hiliteone{89.12} $\pm$ 1.60 & 2.26 \\
				\midrule
				\multirow{3}{*}{\textbf{Kumyk}} & News & 286K &  91.10 $\pm$ 0.86  & 1.53 \\
					& Religion & 227K & 92.47 $\pm$ 1.03  & 1.53 \\\cmidrule{2-5}
				 (r50300) & Average &  & \hiliteone{91.78} $\pm$ 0.94 & 1.53 \\
				\bottomrule
			\end{tabular}\\
		}
			%}
			%%\end{centering}
%		}

@@

		\headerbox{Ongoing and future work}{name=nedostatki,column=3,below=coverage}{
			\begin{itemize}
%				\item case changes for words with one root\\
%					\hiliteone{Ф}инландия \eng{Finland}, \hiliteone{ф}инландиялык \eng{Finnish}
%				\item phonol.\ (vowel harmony, desonorisation) with abbrevs.\\
%					АКШ [акышы] \eng{USA} → АКШ\hiliteone{н}ын / *АКШ\hiliteone{т}ын
%				\item vowel harmony with numbers\\
%					{\qipa 100 [жүз] → 100дүн [жүздүн] / *100нын}
%				\item compound verbs (esp.\ ones with changeable parts)
%				\item gerunds with mono-syllabic V-final verbs\\
%					{\qipa иште- \eng{work} → иштеш / иштөө \eng{working}}\\
%					{\qipa же- \eng{eat} → жеш / *жөө \eng{eating}}
				\item Disambiguation, more stems
				\item Machine translation between these languages
				\item Bring our other Kypchak transducers to comparable performance:\\
				Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
				\item Other Turkic lgs.:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
			\end{itemize}
			%\vspace{0.98em}
		}

		\headerbox{Further information}{name=getting,column=4}{
			\begin{itemize}
				\item Part of Apertium Turkic project:\\
				{\small \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}}
				\item Transducers available live at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
				\item Source code available from apertium's svn repo
				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}

>>>>

		\htwo{Framework: HFST}
		\begin{itemize}
			\item Reimplements Xerox FST formalisms ({\tt lexc} \& {\tt twol})
			\item Also provides a wrapper around popular free/open-source FST toolkits: SFST, OpenFST, and Foma
		\end{itemize}
	}

	\headerbox{Development effort}{name=devel,below=adjectives,column=1}{

		\begin{itemize}
			\item Kumyk transducer based on Kazakh, Tatar transducers
			\item \tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
		\end{itemize}
		\vspace*{-2.5em}
		\hspace*{-2em}\includegraphics[width=0.9\textwidth,angle=270]{../kipchak-lrec-2013/graphs/kumyk-yoldash-cov}
		\vspace*{-2em}
	}

@@

		\headerbox{Categorisation}{name=adjectives,below=exout,column=1}{
		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
			\begin{itemize}
				\item Example:\ morphologically distinct adjective classes (not documented elsewhere)
				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
				\item Our approach:\ categorisation (generates/analyses only correct forms)
			\end{itemize}

@@

			\htwo{Number of stems}
			%\vspace{-1.5em}
			\vspace{-0.5em}
			\begin{center}
			\begin{tabular}{lrrrr}
				\toprule
				\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{4}{c}{\textbf{Number of stems}} \\ \cmidrule{2-5}
				& \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
				\midrule
				Noun & 4582 & 2640 & 2795 & 2568 \\
				Verb & 1193 & 1470 & 1143 & 386 \\
				Adjective & 1211 & 754 & 816 & 219 \\
				Proper noun & 5887 & 5701 & 5361 & 1443 \\
				Adverb & 312 & 171 & 177 & 63 \\
				Numeral & 66 & 63 & 63 & 44 \\
				Conjunction & 77 & 46 & 45 & 13 \\
				Postposition & 50 & 50 & 43 & 12 \\
				Pronoun & 51 & 32 & 28 & 17 \\
				Determiner & 64 & & 39 & 34 & 9 \\
				\midrule
				Total: & \hiliteone{13749} & \hiliteone{11224} & \hiliteone{10737} & \hiliteone{4845} \\
				\bottomrule
			\end{tabular}
			\end{center}
			%\vspace{2em}
			\vspace{-0.5em}

@@

%		\scalebox{0.8}{
%			\begin{tabular}{lll}
%				\toprule
%				\textbf{type} & \textbf{lang} & \textbf{contents} \\ % & \textbf{origin} \\
%				\midrule
%				Encyclopædic & kaz & Wikipedia \\ % wpdump & 20131006 \\
%					& tat & Wikipedia \\ % wpdump & 20130225 \\
%					& kum & — \\ % — & — \\
%				\midrule
%				News & kaz & RFE/RL (azattyq.org) \\ % RFE/RL & azattyq.org 2010 \\
%					& tat & tat.tatar-inform.ru \\ % Татар-информ & tat.tatar-inform.ru 2005-2011 \\
%					& kum & Ёлдаш (yoldash.etnosmi.ru) \\ % Ёлдаш & yoldash.etnosmi.ru \\
%				\midrule
%				Religion & kaz & Quran + Bible \\ % quran + bible & kkitap.net, kuran.kz \\
%					& tat & Quran + New Testament \\ % quran + nt & ibt.org.ru, tanzil.net \\
%					& kum & Genesis + New Testament \\ % genesis + nt & ibt.org.ru \\
%				\bottomrule
%			\end{tabular}
%		}
		\scalebox{0.77}{
			\begin{tabular}{llll}
				\toprule
					& \textbf{Wikipedia} & \textbf{News} & \textbf{Religion} \\
				\midrule
					\texttt{\textbf{Kyrgyz}} & Wikipedia & azattyk.org & Quran + Bible \\
					\texttt{\textbf{Kazakh}} & Wikipedia & azattyq.org & Quran + Bible \\
					\texttt{\textbf{Tatar}} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
					\texttt{\textbf{Kumyk}} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
				\bottomrule
			\end{tabular}

@@

		\scalebox{0.89}{
			\begin{tabular}{llrrr}
				\toprule
				 & \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) & \textbf{Amb.} \\
				\midrule

				\multirow{4}{*}{\textbf{Kyrgyz}} & Wikipedia & M &  $\pm$  &  \\
					&  &  &  $\pm$  &  \\\cmidrule{2-5}
				 (r54474) & Average &  & \hiliteone{} $\pm$  &  \\
				\midrule
				\multirow{4}{*}{\textbf{Kazakh}} & Wikipedia & 25.6M & 85.61 $\pm$ 1.37 & 2.43 \\
					& News & 3.8M & 92.12 $\pm$ 2.72  & 2.88 \\
					& Religion & 851K & 92.49 $\pm$ 1.66  & 2.63 \\\cmidrule{2-5}
				 (r50547) & Average &  & \hiliteone{90.07} $\pm$ 1.91  & 2.64\\
				\midrule
				\multirow{4}{*}{\textbf{Tatar}} & Wikipedia & 159K & 86.35 $\pm$ 2.17  & 2.24\\
					& News & 5.2M & 89.75 $\pm$ 0.07  & 2.30 \\
					& Religion & 382K & 91.25 $\pm$ 2.55 & 2.24 \\\cmidrule{2-5}
				 (r50260) & Average &  & \hiliteone{89.12} $\pm$ 1.60 & 2.26 \\
				\midrule
				\multirow{3}{*}{\textbf{Kumyk}} & News & 286K &  91.10 $\pm$ 0.86  & 1.53 \\
					& Religion & 227K & 92.47 $\pm$ 1.03  & 1.53 \\\cmidrule{2-5}
				 (r50300) & Average &  & \hiliteone{91.78} $\pm$ 0.94 & 1.53 \\
				\bottomrule
			\end{tabular}\\
		}
			%}
			%%\end{centering}
%		}

@@

		\headerbox{Ongoing and future work}{name=nedostatki,column=5}{
			\begin{itemize}
%				\item case changes for words with one root\\
%					\hiliteone{Ф}инландия \eng{Finland}, \hiliteone{ф}инландиялык \eng{Finnish}
%				\item phonol.\ (vowel harmony, desonorisation) with abbrevs.\\
%					АКШ [акышы] \eng{USA} → АКШ\hiliteone{н}ын / *АКШ\hiliteone{т}ын
%				\item vowel harmony with numbers\\
%					{\qipa 100 [жүз] → 100дүн [жүздүн] / *100нын}
%				\item compound verbs (esp.\ ones with changeable parts)
%				\item gerunds with mono-syllabic V-final verbs\\
%					{\qipa иште- \eng{work} → иштеш / иштөө \eng{working}}\\
%					{\qipa же- \eng{eat} → жеш / *жөө \eng{eating}}
				\item Disambiguation, more stems
				\item Clean up Kyrgyz transducer
				\item Machine translation between these languages
				\item Bring our other Kypchak transducers to comparable performance:\\
				Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
				\item Other Turkic lgs.:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
			\end{itemize}
			%\vspace{0.98em}
		}

		\headerbox{Further information}{name=getting,column=0,below=morphtrans}{
			\begin{itemize}
				\item Part of Apertium Turkic project:\\
				{\small \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}}
				\item Transducers available live at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
				\item Source code available from apertium's svn repo
				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
